PapeR Brain

### 2022-01-03 Markdown/PapeR Brain.Rmd
Version 3.3 of Markdown

Hey there! This is an R Markdown document showcasing one particular use of the `networkd3` package. Here, we use the `simpleNetwork()` package to plot a network diagram, used to model biological neural networks.

@@

The graphs you'll see here are a digital iteration of a project I started in 2021, aptly called "Paper Brain." I set up basic neuron models on paper which consists of only two actions: either (1) fire an excitatory pulse when triggered, or (2) fire an inhibitory pulse when triggered.

@@

When we construct the biological manifestations for decision-making operators may be the structural motifs we may want to search for when we consider the expanse of the neural circuits in the human brain.

While this depiction may not be completely generalizable to a brain in its complexity, analyzing the simplest model may present us with an easier model we can use to dissect the true circuits.

@@

### A Variation on the loop

@@

loopNetworkv1 <- data.frame(srcLOOPv1, targetLOOPv1) # construct data frame with source and target vectors

@@

## Limitations

>>>>

Hey there! This is an R Markdown document showcasing one particular use of the `networkd3` package. Here, we use the `simpleNetwork()` package to plot Micro-Networks, used to model biological neural networks in terms of computational functions.

@@

The graphs you'll see here are a digital iteration of a project I started in 2021, aptly called "Paper Brain." I set up basic neuron models on paper which consists of only two actions:

- A neuron will fire an excitatory pulse when triggered
- A neuron will fire an inhibitory pulse when triggered

@@

When we put together the structures for decision-making operators, these may be the structural motifs we may want to search for when we consider the expanse of the neural circuits in the human brain.

While this depiction may not currently be completely generalizable to a brain in its complexity, analyzing the simplest model may present us with an easier model we can use to dissect the true circuits.

@@

### A Loop Variation

The code below constructs a data frame with two extra connections added to both srcLOOPv1 and targetLOOPv1. These extra connections can be seen at the bottom 2 rows of the outputted data frame.

Try to use the nomenclature system to figure out how this modifies the Loop MicroNetwork

@@

# construct data frame with source and target vectors

loopNetworkv1 <- data.frame(srcLOOPv1, targetLOOPv1)

@@

## Combining MicroNetworks

### Isolated MicroNetworks

As mentioned, every neuron which is part of a distinct MicroNetwork has a unique identifier. As a result, multiple networks can be stitched together and viewed as a singular, poly-network assembly. In the following code, we put the "And" and "Loop" networks in the same environment (not neurally connected, however). This is in order to allow you to manipulate both MicroNetworks and visualize the multiplicity of ways that they can be connected to one another.

```{r}
micronetAndLoop <- function() {
  srcBOTH <- c(srcAND, srcLOOP)
  targetBOTH <- c(targetAND, targetLOOP)

  dualComponentNetwork <- data.frame(srcBOTH, targetBOTH)

  simpleNetwork(dualComponentNetwork, fontSize = 14, nodeColour = rainbow(6),
                linkDistance = 100, zoom = TRUE)
}

micronetAndLoop()
```

### Combined MicroNetworks

Manipulating the two networks, they can easily be seen as discrete entities — however, we can also combine these two networks to form a larger, more complex network.

For instance, if we wanted to develop a network model where if A, B, and C were all true, a positive feedback loop would be the output, we can develop the micronetwork here:

```{r}

# The only change made here is adding "Da (Out)" to the end of our src list.

srcAND <- c("Aa (In)", "Ba (In)", "Ca (In)",
            "(1a)+", "(2a)+", "(3a)+",
            "(1a)+", "(2a)+", "(3a)+",
            "(4a)+", "(4a)+", "(6a)-",
            "(7a)-", "(5a)+", "Da (Out)")

# The only change made is adding "Al (In)" to the end of our target list.

targetAND <- c("(1a)+", "(2a)+", "(3a)+",
               "(4a)+", "(4a)+", "(4a)+",
               "(5a)+", "(5a)+", "(5a)+",
               "(6a)-", "(7a)-", "(5a)+",
               "(5a)+", "Da (Out)", "Al (In)")

# The following vectors of neurons were not modified at all.

srcLOOP <- c("Al (In)",
             "(1l)+", "(2l)+", "(3l)+", "(4l)+",
             "(3l)+")

micronetAndLoop()
```

In a case like this, it's as easy as linking the output neuron Da (Out) to the input neuron Al (In), and assembling the rest of the network using the `micronetAndLoop()` function we wrote earlier.

While this neural network appears more complicated, the function reflects exactly what structure we designed for it. In other words, if input neurons Aa (In), Ba (In), and Ca (In) all experience a stimulus, they will result in output neuron Bl (Out) releasing regular pulses due to the positive feedback loop.

We can link up MicroNetworks in many different configurations to model any variety of neuronal pathway, provided that we have a basic understanding of the structural motifs which represent these computational operators in our target organisms.

# Limitations

## Modular Simplicity

(may not account for brain complexity)

## Coloring Limitations

(same color can be confusing, label excitatory / inhibitory / input / output neurons?)

## Modular Generalizability

(not compared to microscopy yet)

## Transmission Directionality

(arrow-paths to show neural transmission)

# Future Directions

One of the primary aims of the PaperBrain / MicroNetwork project is to allow for a simple interface where researchers in Systems or Computational Neuroscience and various other fields can construct these models using language one might ordinarily see while programming. For instance, a programmer could write:

`if (A & B & C) {`
  `while (I == FALSE) {`
    `triggerOutput()`
  `}`
`}`

PapeRBrain could interpret this code as a combination of an And MicroNetwork (recognizing the `&` operator) and a Loop MicroNetwork (recognizing the `while` loop), and construct the neurobiological representation of this code using `networkd3`.

(Looking carefully at this code, the Loop MicroNetwork may be constructed as a Loop unit with some input neuron I acting as an inhibitory element on the positive feedback loop)

Constructing these complex assemblies can be modular, and MicroNetwork units such as "And," "Loop," "Not," and "Or" (among others) can be modularly assembled using the syntax of computer programming (which may be a useful construction interface for programmers).

Another goal of the PBMN project is to inspect the basic modular circuits which form the nervous systems of simpler organisms, and design a directory for MicroNetworks where the relevant units for certain species can be easily retrieved. As a sub-goal, I would like to standardize the nomenclature for neurons in MicroNetworks, to minimize confusion.

Finally, I hope to make the visualizations more functional by designing an interface where the circuit can be interacted with: various neurons can be triggered in order to assess the circuit's redundancies, areas for improvement, or simply effectiveness.
